Convective heat transport in stratified atmospheres at low and high Mach number

### 2016-10-13 tex/nusselt_paper.tex
Minor typo (epsilon, sign error) & spelling fixes.

Atmospheric diffusivities are specified by the Rayleigh number and the Prandtl number.  The
non-dimensional Rayleigh number is
\begin{equation}
\text{Ra} = \frac{g L_z^3 (\Delta S_0 / c_P)}{\nu\chi},
\end{equation}
where $\Delta S_0$ is the entropy difference between the top and bottom of the atmosphere, 
$c_P = R\gamma(\gamma-1)^{-1}$ is the specific heat at constant pressure,
$\nu$ is the kinematic viscosity (viscous diffusivity), and $\chi$ is the thermal dififusivity.  
The relationship between the thermal and viscous diffusivities is
set by the Prandtl number, Pr$ = \nu/\chi$.   We relate the dynamic viscosity, $\mu$, and the thermal conductivity,
$\kappa$, to their corresponding diffusivities such that 
$\nu \equiv \mu/\rho$ and $\chi \equiv \kappa/\rho$.  As a result, $\text{Ra} \propto (\nu\chi)^{-1} \propto
\rho^2$.  The atmospheres studied here with $n_{\rho} = 3$ experience an increase in the Rayleigh number 
by a factor of 400 across the domain.  This formulation leaves Pr
constant throughout the depth of the atmosphere, and in this letter we impose $\text{Pr} = 1$.
Throughout this letter, we specify Ra at the top of the atmosphere.

@@

The atmosphere is contained between two impenetrable, stress free, fixed temperature boundaries at
the top and bottom of the domain such that $w = \partial_z u = T_1 = 0$ at $z = \{0, L_z\}$. The domain
is periodic in the horizontal. We utilize the novel Dedalus\footnote{http://dedalus-project.org/} pseudospectral framework 
 to time-evolve Eqs. 
\ref{eqn:continuity_eqn}-\ref{eqn:energy_eqn} using an implicit-explicit third-order four-step 
Runge-Kutta timestepping scheme \cite{ascher&all1997}.  
Variables are time-evolved on a dealiased Chebyshev (vertical)
and Fourier (horizontal) domain in which the
physical grid dimensions are 3/2 the size of the coefficient grid.  Our physical grid sizes range from
96x384 grid points at the lowest values of Ra to 1152x4608 grid points at Ra $\geq 10^{7}$. 
By using IMEX timestepping, we can implicity step the stiff linear acoustic wave contribution and we are able to
efficiently study flows at moderate ($Ma \approx 1$) and very low ($Ma \approx 10^{-4}$)
Mach number (Fig. \ref{fig:ma_v_eps}b).

@@

\begin{figure}[t]
\includegraphics[width=3.4375in]{./figs/fluxes_fig.png}
\caption{Time-averaged flux profiles for high (top) and low (bottom) Mach number flows at $Ra = 10^6$.  
The dashed lines correpond to the
enthalpy flux (orange, positive) and kinetic energy flux (purple, negative).  The grey dash-dot line is the
viscous flux, the blue dash-dot-dot line is the potential energy flux, 
and the green dotted line is the radiative flux with the adiabatic radiative flux removed. All
fluxes are normalized by $F_{ref} - F_A$, as in the calculation of the Nusselt number.  The solid black line is
the properly normalized sum of all the fluxes, and under this normalization its height-averaged value is the
Nusselt number.
\label{fig:flux_profiles} }
\end{figure}

At high values of Ra, the heat transport properties of the systems become increasingly complex and time-dependent.
Large $\epsilon$ flows exhibit two local maxima in the enthalpy flux and kinetic energy flux: one in the upper 
atmosphere caused by the shocks, and one in the lower atmosphere caused by the deep mixing of convective motions
(see Fig. \ref{fig:flux_profiles}a).
At low $\epsilon$, only the deep maximum is present (Fig. \ref{fig:flux_profiles}b).  
Furthermore, the presence
of fixed-temperature boundary conditions allows the flux at the boundaries to vary.  Many runs at $Ra > 10^5$ and
$\epsilon = 10^{-4}$ exhibit time-dependent states of stratified convection (such as those shown in 
Fig. \ref{fig:entropy_snapshots}b\&c), in which the flux entering the system at the bottom of the atmosphere exceeds
that at the top.  These states are punctuated by states of vigorous shearing, similar to those previouslyl
reported in two-dimensional RB convection \cite{goluskin&all2014}.  These shearing states flatten the bottom temperature
gradient towards adiabatic, allowing the excess energy to exit through the upper boundary.  These shearing
states will be covered in more detail in a future paper.  Regardless, a proper long-term average over shearing
and non-shearing states retrieves an invariant flux profile (and therefore Nu profile) throughout the depth
of the atmosphere.

@@

The dynamics of these polytropic solutions are complex and highly time-dependent, even in two dimensions.
Time-dependent oscillating shear states have developed spontaneously, as seen before in RB convection
\cite{goluskin&all2014}.  While computationally difficult, the highest values of Ra and the lowest value
of $\epsilon$ studied here are far below the values found in nature.  If the scalings of Nu and Ma
presented here hold, then under solar conditions ($\text{Ra }\approx 10^{20}$, $\text{Ma }\approx 10^{-4}$), we expect that $\epsilon \approx 10^{20}$ and
$\text{Nu }\approx 10^{6}$.

>>>>

Atmospheric diffusivities are specified by the Rayleigh number and the Prandtl number.  The
non-dimensional Rayleigh number is
\begin{equation}
\text{Ra} = \frac{g L_z^3 (\Delta S_0 / c_P)}{\nu\chi},
\end{equation}
where $\Delta S_0$ is the entropy difference between the top and bottom of the atmosphere, 
$c_P = R\gamma(\gamma-1)^{-1}$ is the specific heat at constant pressure,
$\nu$ is the kinematic viscosity (viscous diffusivity), and $\chi$ is the thermal diffusivity.  
The relationship between the thermal and viscous diffusivities is
set by the Prandtl number, Pr$ = \nu/\chi$.   We relate the dynamic viscosity, $\mu$, and the thermal conductivity,
$\kappa$, to their corresponding diffusivities such that 
$\nu \equiv \mu/\rho$ and $\chi \equiv \kappa/\rho$.  As a result, $\text{Ra} \propto (\nu\chi)^{-1} \propto
\rho^2$.  The atmospheres studied here with $n_{\rho} = 3$ experience an increase in the Rayleigh number 
by a factor of 400 across the domain.  This formulation leaves Pr
constant throughout the depth of the atmosphere, and in this letter we impose $\text{Pr} = 1$.
Throughout this letter, we specify Ra at the top of the atmosphere.

@@

The atmosphere is contained between two impenetrable, stress free, fixed temperature boundaries at
the top and bottom of the domain such that $w = \partial_z u = T_1 = 0$ at $z = \{0, L_z\}$. The domain
is periodic in the horizontal. We utilize the novel Dedalus\footnote{http://dedalus-project.org/} pseudospectral framework 
 to time-evolve Eqs. 
\ref{eqn:continuity_eqn}-\ref{eqn:energy_eqn} using an implicit-explicit third-order four-step 
Runge-Kutta timestepping scheme \cite{ascher&all1997}.  
Variables are time-evolved on a dealiased Chebyshev (vertical)
and Fourier (horizontal) domain in which the
physical grid dimensions are 3/2 the size of the coefficient grid.  Our physical grid sizes range from
96x384 grid points at the lowest values of Ra to 1152x4608 grid points at Ra $\geq 10^{7}$. 
By using IMEX timestepping, we can implicitly step the stiff linear acoustic wave contribution and we are able to
efficiently study flows at moderate ($Ma \approx 1$) and very low ($Ma \approx 10^{-4}$)
Mach number (Fig. \ref{fig:ma_v_eps}b).

@@

\begin{figure}[t]
\includegraphics[width=3.4375in]{./figs/fluxes_fig.png}
\caption{Time-averaged flux profiles for high (top) and low (bottom) Mach number flows at $Ra = 10^6$.  
The dashed lines correspond to the
enthalpy flux (orange, positive) and kinetic energy flux (purple, negative).  The grey dash-dot line is the
viscous flux, the blue dash-dot-dot line is the potential energy flux, 
and the green dotted line is the radiative flux with the adiabatic radiative flux removed. All
fluxes are normalized by $F_{ref} - F_A$, as in the calculation of the Nusselt number.  The solid black line is
the properly normalized sum of all the fluxes, and under this normalization its height-averaged value is the
Nusselt number.
\label{fig:flux_profiles} }
\end{figure}

At high values of Ra, the heat transport properties of the systems become increasingly complex and time-dependent.
Large $\epsilon$ flows exhibit two local maxima in the enthalpy flux and kinetic energy flux: one in the upper 
atmosphere caused by the shocks, and one in the lower atmosphere caused by the deep mixing of convective motions
(see Fig. \ref{fig:flux_profiles}a).
At low $\epsilon$, only the deep maximum is present (Fig. \ref{fig:flux_profiles}b).  
Furthermore, the presence
of fixed-temperature boundary conditions allows the flux at the boundaries to vary.  Many runs at $Ra > 10^5$ and
$\epsilon = 10^{-4}$ exhibit time-dependent states of stratified convection (such as those shown in 
Fig. \ref{fig:entropy_snapshots}b\&c), in which the flux entering the system at the bottom of the atmosphere exceeds
that at the top.  These states are punctuated by states of vigorous shearing, similar to those previously
reported in two-dimensional RB convection \cite{goluskin&all2014}.  These shearing states flatten the bottom temperature
gradient towards adiabatic, allowing the excess energy to exit through the upper boundary.  These shearing
states will be covered in more detail in a future paper.  Regardless, a proper long-term average over shearing
and non-shearing states retrieves an invariant flux profile (and therefore Nu profile) throughout the depth
of the atmosphere.

@@

The dynamics of these polytropic solutions are complex and highly time-dependent, even in two dimensions.
Time-dependent oscillating shear states have developed spontaneously, as seen before in RB convection
\cite{goluskin&all2014}.  While computationally difficult, the highest values of Ra and the lowest value
of $\epsilon$ studied here are far below the values found in nature.  If the scalings of Nu and Ma
presented here hold, then under solar conditions ($\text{Ra }\approx 10^{20}$, $\text{Ma }\approx 10^{-4}$), we expect that $\epsilon \approx 10^{-20}$ and
$\text{Nu }\approx 10^{6}$.
